---
title: "ReprRes_2"
author: "B. van Bekkum"
date: "7 januari 2018"
output: html_document
---

#Loading and preprocessing the data
```{r echo=TRUE, warning=FALSE, message=FALSE}

setwd("~/Coursera/5_RerpRes/Data")
library(dplyr)
library(ggplot2)

myset<- read.csv("activity.csv", colClasses = "character")
myset$steps <- as.numeric(myset$steps)
myset$date <- as.Date(myset$date, format = "%Y-%m-%d")
myset$interval <- as.numeric(myset$interval)

```


#What is mean total number of steps taken per day?

```{r echo=TRUE, warning=FALSE}
stpsday <- myset %>% group_by(date) %>% summarise(totalsteps = sum(steps), na.rm = TRUE)
hist(stpsday$totalsteps, main = "", col = "blue", xlab = "Total steps per day", breaks = 20)
```


- The mean steps per day is: `r round(mean(stpsday$totalsteps, na.rm = TRUE), digits = 2)`
- The median steps per day is: `r median(stpsday$totalsteps, na.rm = TRUE)`


#What is the average daily activity pattern?


```{r}
stpsint <- myset %>% group_by(interval) %>% summarise(meansteps = mean(steps, na.rm = TRUE))
plot(stpsint, type = "l", main = "Mean steps per interval")
```

- The interval with the highest mean is: `r  stpsint[stpsint$meansteps == max(stpsint$meansteps),1]`.
- The total number of rows is: `r nrow(myset)`. The number of rows with missing values is: `r sum(is.na(myset))`.

#Imputing missing values

```{r}
myset2 <- myset
for (i in 1:nrow(myset2)){
    if (is.na(myset2$steps[i])){ 
      myset2$steps[i] <- stpsint$meansteps[i]}
}

stpsday2 <- myset2 %>% group_by(date) %>% summarise(totalsteps = sum(steps))
hist(stpsday2$totalsteps, main = "Total steps a day, na's replaced with interval mean", 
     col = "gray", xlab = "Total steps per day", breaks = 20)
```

- The mean steps per day is: `r mean(stpsday2$totalsteps, na.rm = TRUE)`
- The median steps per day is: `r median(stpsday2$totalsteps, na.rm = TRUE)`


#Are there differences in activity patterns between weekdays and weekends?

```{r}
##myset$test <- weekdays(myset$date, abbreviate = T)
myset$date <- as.character(myset$date)
myset$date <- strptime(myset$date, format = "%Y-%m-%d")
wknd <- c("za", "zo")
myset$weekday <- as.factor(with(myset, ifelse(weekdays(date, abbreviate = T) %in% wknd, TRUE, FALSE)))
levels(myset$weekday) <- c("weekday", "weekend")
mysub <- myset[,!names(myset) == 'date']
```

```{r}
mysetwk <- mysub %>% group_by(weekday, interval) %>% summarise(meansteps = mean(steps, na.rm = TRUE))
qplot(interval, meansteps, data = mysetwk, facets = .~weekday, geom = c("line"),
      main = "Mean steps per interval", ylab = "Mean number of steps", xlab = "Interval")

```